Per capita consumption

### 2023-12-14 PerCapitaConsumption.Rmd
knits
WIP evaluate consumption estimates

@@

@@

This is the function code

@@

This is the current result

@@

biomass_eaten <- calc_pred_cons(eat = eat, grazing = grazing, vol = vol, biolprm = biol, runprm = runprm)

@@

>>>>

# this hardcodes a d.name in a different folder because I don't want multiple copies

if(!dir.exists(d.name)){
  d.name <- "/Users/sarah.gaichas/Documents/0_Data/ms-keyrun/simulated-data/atlantisoutput/NOBA_march_2020"
}

@@

  # look directly at file
  prod.nc = ncdf4::nc_open(file.path(dir, nc_prod))

  names(prod.nc$var)

@@

This is the function code. Consumption still seems low so lets not exlude sediment layer.

```{r}
calc_pred_cons <- function(eat, grazing, vol, fgs, biolprm, runprm){

  # Conversion factor from mg N to tonnes wet-weight
  bio_conv <- biolprm$redfieldcn * biolprm$kgw2d / 1000000000

  # Eat and grazing are per species, agecl, polygon, and time.
  # Therefore, we need to aggregate the vol over layers.
  # Eat calculated in atecology/atprocess.C
  # Eat( ) is defined here : \trunk\atlantis\atecology\atprocess.c(644)

  # SKG: The main question is the UNITS of eat and grazing.
  # manual says this is either mg/m3/day or mg/day/individ
  # calculations here assume mg/m3/day so multiplying by box vol
  # is this m3 assuming the entire box volume or only where groups feed?
  # do we need to match species to layers to get appropriate volume?
  # VERTnight_XXX and VERTday_XXX in biol.prm

  # biomass pool groups may get need volume expansion while
  # age structured are per capita by age group?

  # all layers? even sediment? try filtering out layer > 7 to remove sediment volume
  # WARNING: ensure that layer 7 (processed via atlantisom) isn't unique to NOBA
  # layer 7 is unique to NOBA, use max of layers as sediment layer
  # however, unclear this is necessary, try without

  colnames(fgs) <- tolower(colnames(fgs))

  # check for GroupType or InvertType
  if (!("grouptype" %in% colnames(fgs) | "inverttype"%in% colnames(fgs))) {
    stop(paste("The columns GroupType or InvertType ars not in your functional\n",
               "groups file."))
  }

  # change inverttype to grouptype, contents should be the same
  if("inverttype" %in% names(fgs)) names(fgs)[names(fgs) == 'inverttype'] <- 'grouptype'

  fgs$grouptype <- tolower(fgs$grouptype)

  # is there a different calculation for benthos? don't expand to volume?
  # not using this yet
  pooltype <- fgs |>
    dplyr::select(species=name, grouptype) |>
    dplyr::mutate(pooltype = dplyr::case_when((grouptype %in% c("lg_inf",
                                                                "sm_inf",
                                                                "sed_bact",
                                                                "sed_ep_ff",
                                                                "sed_ep_other",
                                                                "mob_ep_other",
                                                                "coral",
                                                                "sponge")) ~ "benthos",
                                              (grouptype %in% c("lg_phy")) ~ "lg_phy",
                                              TRUE ~ "alllayers"))
  # not using this either
  sedlayer <- max(vol$layer)

  vol <- aggregate(atoutput ~ polygon + time, data = vol, sum)
  colnames(vol)[colnames(vol) == "atoutput"] <- "vol"

  # Combine eat and grazing! Calculate eaten biomass
  biomass_eaten <- rbind(eat, grazing)
  biomass_eaten <- merge(biomass_eaten, vol,
    by = c("time", "polygon"))

  # todo: previously AK noted that the bio_eaten values were too high
  # the new values need to be checked to see if they give realistic
  # results
  # removed layer 7 in volume which should expand mg N to only water?
  # total consumption and per capita from this look ok
  # snapshot so should be daily value?

  biomass_eaten$bio_eaten <- with(biomass_eaten,
    atoutput * vol * bio_conv)

  biomass_eaten$atoutput <- biomass_eaten$bio_eaten

  return(biomass_eaten)
}

@@

This is the current result compared with all layers

@@

biomass_eaten <- calc_pred_cons(eat = eat, grazing = grazing, vol = vol, fgs = fgs, biolprm = biol, runprm = runprm)

biomass_eaten_current <- atlantisom::calc_pred_cons(eat = eat, grazing = grazing, vol = vol, biolprm = biol, runprm = runprm)

@@

Keeping the sediment layer in makes a small difference.

```
> head(biomass_eaten_current)
  time polygon        species agecl    atoutput          vol   bio_eaten
1    1       1 Long_rough_dab     1   0.6127621 2.865986e+13   0.6127621
2    1       1 Long_rough_dab     8  20.7267353 2.865986e+13  20.7267353
3    1       1 Long_rough_dab    10   6.1992620 2.865986e+13   6.1992620
4    1       1        Haddock     2  77.4729324 2.865986e+13  77.4729324
5    1       1         Saithe     9   9.1759148 2.865986e+13   9.1759148
6    1       1         Saithe     3 173.4369707 2.865986e+13 173.4369707
> head(biomass_eaten)
  time polygon        species agecl    atoutput          vol   bio_eaten
1    1       1 Long_rough_dab     1   0.6148762 2.875874e+13   0.6148762
2    1       1 Long_rough_dab     8  20.7982434 2.875874e+13  20.7982434
3    1       1 Long_rough_dab    10   6.2206497 2.875874e+13   6.2206497
4    1       1        Haddock     2  77.7402172 2.875874e+13  77.7402172
5    1       1         Saithe     9   9.2075721 2.875874e+13   9.2075721
6    1       1         Saithe     3 174.0353353 2.875874e+13 174.0353353
```
